---
title: "Project1"
author: "Erica Laughlin"
date: "July 28, 2017"
output: html_document
---

First I need to load required packages, then download, unzip, and read the data.

```{r}
library(knitr)
library(dplyr)
library(ggplot2)
library(downloader)
library(tidyr)
library(mice)

download("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", dest="activity.zip", mode = "wb")
unzip("activity.zip")
activity <- read.csv("activity.csv")  
```

Exploring the data a little.

```{r}
dim(activity)  
head(activity)  
summary(activity)
```

Generating a histogram of total daily steps taken.

```{r}
activity$date <- as.factor(activity$date)
stepsbydate <- group_by(activity, date) %>% 
  summarize(total_steps = sum(steps))
hist(stepsbydate$total_steps, main = "Total Daily Steps", xlab = "Total Daily Steps", col = "green")
```

The histogram shows us that our subject most frequently took between 10,000 and 15,000 steps in a day.  Calculating the mean and median measurements for total daily steps.

```{r}
mean(stepsbydate$total_steps, na.rm = TRUE)
median(stepsbydate$total_steps, na.rm = TRUE)
```

This validates our earlier finding.  The average number of steps taken in a day is about 10,766.  Next I'm going to create a version of the dataframe with missing values removed, and then generate a time series plot of the average number of steps taken.

```{r}
activity$interval <- as.factor(activity$interval)
activity2 <- na.omit(activity)
stepsbyint <- group_by(activity2, interval) %>% 
  summarize(mean_steps = mean(steps))
plot(stepsbyint)
lines(stepsbyint, lwd = 2, col = "red")
```

There is a tall peak in the data that occurs sometime in the morning.  Now I'm going to determine which exact 5-minute interval has the highest average number of steps taken.

```{r}
highestmean <- max(stepsbyint$mean_steps)
subset(stepsbyint, mean_steps == highestmean)
```

Looks like, on average, the highest average number of steps taken is at 8:35 AM each day, with an average of about 206 steps taken.  Now, I'm going to create a version of the data frame that imputes missing data with predictive mean matching using the mice package.

```{r}
ini <- mice(activity, maxit=0, pri=F)
pred <- ini$pred
pred[, "date"] <- 0
imp <- mice(activity, m=1, maxit=1, pred=pred)
imp
```

Let's take a look at the first 100 lines of the new dataset.

```{r}
activity3 <- complete(imp)
head(activity3, n=100)
```

Generating a new histogram of total daily steps taken, now with imputed values.  Calculating mean and median.

```{r}
stepsbydate2 <- group_by(activity3, date) %>% 
  summarize(total_steps = sum(steps))
hist(stepsbydate2$total_steps, main = "Total Daily Steps - Imputed", xlab = "Total Daily Steps", col = "green")
mean(stepsbydate2$total_steps, na.rm = TRUE)
median(stepsbydate2$total_steps, na.rm = TRUE)
```

Notice the shape of the graph is the same, but the frequencies are higher.  The new mean is approximately 10,665 steps (slightly lower than the original calculation of 10,766) and the new median is 10,600 (versus the original of 10,765).  Now let's look at the difference between weekdays and weekends.

```{r}
activity3$date <- as.Date(activity3$date)
activity3$weekday <- weekdays(activity3$date)
wkend <- c("Saturday", "Sunday")
wkday <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
wkend_activity <- subset(activity3, weekday == wkend)
wkday_activity <- subset(activity3, weekday == wkday)
```

